Supplementary Material for Asymmetry paper

### 2020-07-27 Supplementary/Sup_mat.Rmd
New pdf version

---
title: "Supplementary Material for Asymmetry paper"
output: 
  bookdown::html_document2:
    theme: journal
    toc: true
bibliography: ["Literature.bib"]
biblio-style: "apalike"
link-citations: true
---

@@

Probability matching may be due in part to a misunderstanding of probability [@REIMERS201811], but it has also been argued to reﬂect a reasonable tendency to seek out and exploit patterns in sequential events [@gaissmaier2008smart; @wolford2004searching]. @Gao2015 argue that when a person selects the more likely option on every trial (a strategy known as _Maximising_), they are accepting a loss for a minority of trials (20% in this example). As demonstrated in @KahnemanChoicesValuseFrames, people tend to ﬁnd certain loss very aversive, so a strategy that includes a certain loss may be disregarded as a “solution” to the problem at hand [@Goodnow1955]. The only way to avoid this loss is to make an attempt to detect and exploit any potential pattern. In the context of deciding where to ﬁxate between two potential targets locations, in both @morvan2012human and @clarke2015failure, the probability for each location to contain the target target was 50%. Given the philosophical and experimental arguments described above, participants may react to this choice between equally likely options with idiosyncratic pattern-seeking behavior [@gaissmaier2008smart; @yellott1969probability; @wolford2004searching]. In @clarke2015failure, this may have interfered with participants using information about their own ability to make decisions. Instead, they may have been attempting to ﬁgure out patterns in the sequence of targets to make better guesses about which one was likely to be the target on each trial. Adding in a bias for one option to be more likely may cater to people’s tendency to seek out, and attempt to exploit, patterns. It also breaks the Buridan’s Ass deadlock and makes the decision about which of two goals to attempt easier. In this Experiment, unequal probability was introduced into the paradigm used by @clarke2015failure in order to investigate 1) whether participants would make use of probability information in deciding where to ﬁxate, and 2) whether probability manipulations might facilitate the decision about whether to ﬁxate between two potential target locations or to look directly at one of them. If so, it would suggest that at least part of the reason for the poor decisions observed previously was the (unnatural) balance between the two potential targets. Additionally,this experiment will provide a useful replication of both the ﬁxations decision results [@morvan2012human; @clarke2015failure] and the probability matching tendency [@Vulkan2000ProbMatching] in a new context.

@@

In the first session, participants were instructed to remain ﬁxated at the center and identify a letter that would appear in one of two boxes. At the start of each trial,participantsﬁxatedacentral black cross on a grayscale background and pressed the spacebar. The cross was ﬂanked by two boxes which were a lighter shade of grey and occupied 1◦ of visual angle. After a stable ﬁxation had been maintained for 700ms, the target would appear in one of the two boxes for 500ms. The target was a white letter, 0.4&deg; of visual angle, drawn using the Sloan font as these letters are generally of equal recognisability at different viewing angles[@sloan1952comparison]. Ten letters were used and one was selected randomly on each trial to be the target letter. Participants were then presented with a screen prompting them to report which letter was presented by clicking on the corresponding character. An illustration of each trial in Session 1 can be seen in Figure \@ref(fig:ProbSession1).

@@

The boxes were presented on either side of the ﬁxation cross at several different eccentricities (3.1&deg;,4.3&deg;,5.8&deg;,7.5&deg;,9.3&deg;,11.1&deg;,12.5&deg;,& 13.7&deg;). Each eccentricity was repeated 12 times in a row before moving on to another eccentricity (the order of these sets of 12 eccentricities was random),for a total of 96 trials in a block.

After each block, participants were offered a break before recalibrating and moving onto the next block. Participants completed four of these blocks. The data from this ﬁrst session were used to tailor the separations that would be used for the second session. A “switch-point” was calculated for each participant based on their Session 1 performance. This was then used as an anchor point from which to calculate the other separations to be used in Session 2. The switch-point for each participant was the separation at which the participant was 68.5% accurate. The accuracy level of 68.5% is the mid-point between 55% and 82%; these two values are the points at which participants should switch from ﬁxating the central box to the side box in the Symmetric and Bias conditions, respectively(see below for more details).

@@

From this point, 6 other separations were calculated $\pm$ 1&deg;, 2&deg;, and 3&deg;. There were two other points included which acted as anchors, one being a very large distance (19.4&deg;), the other very small (1.9&deg;). In Session 2, participants started by ﬁxating a cross that appeared above where the boxes would appear. The cross would appear at the midpoint between the centre box and either the left or right box with an equal probability. Once they had ﬁxated the cross, they were instructed to press the spacebar. After a stable ﬁxation was detected for 700ms, three boxes would be presented (Figure \@ref(fig:ProbSession2)). One box would appear in the centre with the other two spaced equally on either side with separations calculated as above, based on Session 1 performance for that particular participant. Once these boxes had appeared, participants were instructed to ﬁxate one of the three boxes. Participants were told that the target would never appear in the central box, however they could choose to ﬁxate this location. After they had ﬁxated one of the boxes, the letter appeared in one of the side boxes for 500ms. After this, the 10 letter stimuli were presented on screen for them to select which letter they had seen. Each separation appeared 10 times within a block, for a total of 90 trials. The order of separations was randomised (rather than blocked, as it was in Session 1).

@@

The optimal strategy is to ﬁxate whichever of the three boxes maximises expected accuracy. We refer to a participant’s expected accuracy if they had followed the optimal strategy, as “optimal accuracy”.  Using the psychometric curves ﬁt to each participant’s Session 1 data, we can calculate how likely a given person is to detect the target at various distances. This was then used to predict how accurate that same person would be if they had ﬁxated one of the side boxes, or the centre box, by using their average accuracy for a given distance from the left and right box ($Bl$ and $Br$ respectively) in Session 1. The ﬁxation choice that gave the greatest expected accuracy was then selected as the optimal decision for the respective box separation. As probability was also a factor in this experiment, this had to be accounted for by multiplying the chance of detecting the target in either box by the probability that the box would contain the target ($Pl$ and $Pr$). The formula for expected accuracy is therefore: $(Bl \times Pl) + (Br \times Pr)$. Expected accuracy given optimal ﬁxations differs between the Symmetric and Biased conditions. For example, in the Symmetric condition, participants could expect a 55% success rate if they ﬁxated the optimal location when the targets were far apart. This value comes from assuming they would be $\approx$ 100% accurate for the ﬁxated box, and at chance level for the non-ﬁxated box (10%). This gives us $(1 \times 0.5) + (0.1 \times 0.5) = 55%$. To get the lower limit for the Bias condition, we simply change the chance for each box to contain the target (additionally, we assume that the participant would ﬁxate the most likely box). This gives us $(1 \times 0.8) + (0.1 \times 0.2) = 82%$. This same formula can be used to calculate expected accuracy had the participant ﬁxated the central box. In this case, $Bl = Br$, and so would simply be put into the formula using the appropriate values for $P$, as demonstrated above.

@@

The choices made by participants on where to look are summarized in Figure \@ref(fig:ProbProp). The optimal strategy in this situation is to ﬁxate the central box when the two sides boxes are near (i.e., when their distance from centre, $\Delta$, is closer than each participant’s switch point). When they are far (i.e. delta is greater than the switch point), a participant behaving optimally should ﬁxate either of the side boxes when in the symmetric condition, and the most-likely box when in the biased condition. The ﬁrst panel of Figure \@ref(fig:ProbProp) clearly shows that our participants did not follow this strategy.  However, while adding a bias to the location of the target did not help participants to behave optimally, we can see that it did have an effect on their behavior, as they were much less likely to ﬁxate the central box. Furthermore, when ﬁxating one of the two side squares, they ﬁxated the square that was most likely to contain the target (Figure \@ref(fig:ProbProp), central and right panel) almost all the time

@@

We can also look at how the choice of where to look inﬂuenced participants’ accuracy. We use the target detection models from Session 1 to calculate the expected accuracy given either (i) an optimal, (ii) counter-optimal, or (iii) the observed strategy. This measure eliminates the variance present in the actual accuracy data due to the random location of the target and participant’s guessing the correct answer by chance. These data are summarised in Figure \@ref(fig:ProbExpAcc). We can see that in the symmetric condition, with the exception of one individual, our participants behaved in a way that gave rise to expected accuracies much closer to the counter-optimal strategy than optimal. Expected accuracy was higher in the biased condition, with four participants reaching optimal performance, with the rest ending up somewhere between the optimal and counter-optimal.

>>>>

---
title: "Supplementary Material for Asymmetry paper"
output: 
  bookdown::pdf_document2:
    #theme: journal
    toc: true
bibliography: ["Literature.bib"]
biblio-style: "apalike"
link-citations: true
---

@@

Probability matching may be due in part to a misunderstanding of probability [@REIMERS201811], but it has also been argued to reflect a reasonable tendency to seek out and exploit patterns in sequential events [@gaissmaier2008smart; @wolford2004searching]. @Gao2015 argue that when a person selects the more likely option on every trial (a strategy known as _Maximising_), they are accepting a loss for a minority of trials (20% in this example). As demonstrated in @KahnemanChoicesValuseFrames, people tend to find certain loss very aversive, so a strategy that includes a certain loss may be disregarded as a “solution” to the problem at hand [@Goodnow1955]. The only way to avoid this loss is to make an attempt to detect and exploit any potential pattern. In the context of deciding where to fixate between two potential targets locations, in both @morvan2012human and @clarke2015failure, the probability for each location to contain the target target was 50%. Given the philosophical and experimental arguments described above, participants may react to this choice between equally likely options with idiosyncratic pattern-seeking behavior [@gaissmaier2008smart; @yellott1969probability; @wolford2004searching]. In @clarke2015failure, this may have interfered with participants using information about their own ability to make decisions. Instead, they may have been attempting to figure out patterns in the sequence of targets to make better guesses about which one was likely to be the target on each trial. Adding in a bias for one option to be more likely may cater to people’s tendency to seek out, and attempt to exploit, patterns. It also breaks the Buridan’s Ass deadlock and makes the decision about which of two goals to attempt easier. In this Experiment, unequal probability was introduced into the paradigm used by @clarke2015failure in order to investigate 1) whether participants would make use of probability information in deciding where to fixate, and 2) whether probability manipulations might facilitate the decision about whether to fixate between two potential target locations or to look directly at one of them. If so, it would suggest that at least part of the reason for the poor decisions observed previously was the (unnatural) balance between the two potential targets. Additionally,this experiment will provide a useful replication of both the fixations decision results [@morvan2012human; @clarke2015failure] and the probability matching tendency [@Vulkan2000ProbMatching] in a new context.

@@

In the first session, participants were instructed to remain fixated at the center and identify a letter that would appear in one of two boxes. At the start of each trial,participants fixated a central black cross on a grayscale background and pressed the spacebar. The cross was flanked by two boxes which were a lighter shade of grey and occupied 1◦ of visual angle. After a stable fixation had been maintained for 700ms, the target would appear in one of the two boxes for 500ms. The target was a white letter, 0.4&deg; of visual angle, drawn using the Sloan font as these letters are generally of equal recognisability at different viewing angles[@sloan1952comparison]. Ten letters were used and one was selected randomly on each trial to be the target letter. Participants were then presented with a screen prompting them to report which letter was presented by clicking on the corresponding character. An illustration of each trial in Session 1 can be seen in Figure \@ref(fig:ProbSession1).

@@

The boxes were presented on either side of the fixation cross at several different eccentricities (3.1&deg;,4.3&deg;,5.8&deg;,7.5&deg;,9.3&deg;,11.1&deg;,12.5&deg;,& 13.7&deg;). Each eccentricity was repeated 12 times in a row before moving on to another eccentricity (the order of these sets of 12 eccentricities was random),for a total of 96 trials in a block.

After each block, participants were offered a break before recalibrating and moving onto the next block. Participants completed four of these blocks. The data from this first session were used to tailor the separations that would be used for the second session. A “switch-point” was calculated for each participant based on their Session 1 performance. This was then used as an anchor point from which to calculate the other separations to be used in Session 2. The switch-point for each participant was the separation at which the participant was 68.5% accurate. The accuracy level of 68.5% is the mid-point between 55% and 82%; these two values are the points at which participants should switch from fixating the central box to the side box in the Symmetric and Bias conditions, respectively(see below for more details).

@@

From this point, 6 other separations were calculated $\pm$ 1&deg;, 2&deg;, and 3&deg;. There were two other points included which acted as anchors, one being a very large distance (19.4&deg;), the other very small (1.9&deg;). In Session 2, participants started by fixating a cross that appeared above where the boxes would appear. The cross would appear at the midpoint between the centre box and either the left or right box with an equal probability. Once they had fixated the cross, they were instructed to press the spacebar. After a stable fixation was detected for 700ms, three boxes would be presented (Figure \@ref(fig:ProbSession2)). One box would appear in the centre with the other two spaced equally on either side with separations calculated as above, based on Session 1 performance for that particular participant. Once these boxes had appeared, participants were instructed to fixate one of the three boxes. Participants were told that the target would never appear in the central box, however they could choose to fixate this location. After they had fixated one of the boxes, the letter appeared in one of the side boxes for 500ms. After this, the 10 letter stimuli were presented on screen for them to select which letter they had seen. Each separation appeared 10 times within a block, for a total of 90 trials. The order of separations was randomised (rather than blocked, as it was in Session 1).

@@

The optimal strategy is to fixate whichever of the three boxes maximises expected accuracy. We refer to a participant’s expected accuracy if they had followed the optimal strategy, as “optimal accuracy”.  Using the psychometric curves fit to each participant’s Session 1 data, we can calculate how likely a given person is to detect the target at various distances. This was then used to predict how accurate that same person would be if they had fixated one of the side boxes, or the centre box, by using their average accuracy for a given distance from the left and right box ($Bl$ and $Br$ respectively) in Session 1. The fixation choice that gave the greatest expected accuracy was then selected as the optimal decision for the respective box separation. As probability was also a factor in this experiment, this had to be accounted for by multiplying the chance of detecting the target in either box by the probability that the box would contain the target ($Pl$ and $Pr$). The formula for expected accuracy is therefore: $(Bl \times Pl) + (Br \times Pr)$. Expected accuracy given optimal fixations differs between the Symmetric and Biased conditions. For example, in the Symmetric condition, participants could expect a 55% success rate if they fixated the optimal location when the targets were far apart. This value comes from assuming they would be $\approx$ 100% accurate for the fixated box, and at chance level for the non-fixated box (10%). This gives us $(1 \times 0.5) + (0.1 \times 0.5) = 55%$. To get the lower limit for the Bias condition, we simply change the chance for each box to contain the target (additionally, we assume that the participant would fixate the most likely box). This gives us $(1 \times 0.8) + (0.1 \times 0.2) = 82%$. This same formula can be used to calculate expected accuracy had the participant fixated the central box. In this case, $Bl = Br$, and so would simply be put into the formula using the appropriate values for $P$, as demonstrated above.

@@

The choices made by participants on where to look are summarized in Figure \@ref(fig:ProbProp). The optimal strategy in this situation is to fixate the central box when the two sides boxes are near (i.e., when their distance from centre, $\Delta$, is closer than each participant’s switch point). When they are far (i.e. delta is greater than the switch point), a participant behaving optimally should fixate either of the side boxes when in the symmetric condition, and the most-likely box when in the biased condition. The first panel of Figure \@ref(fig:ProbProp) clearly shows that our participants did not follow this strategy.  However, while adding a bias to the location of the target did not help participants to behave optimally, we can see that it did have an effect on their behavior, as they were much less likely to fixate the central box. Furthermore, when fixating one of the two side squares, they fixated the square that was most likely to contain the target (Figure \@ref(fig:ProbProp), central and right panel) almost all the time

@@

We can also look at how the choice of where to look influenced participants’ accuracy. We use the target detection models from Session 1 to calculate the expected accuracy given either (i) an optimal, (ii) counter-optimal, or (iii) the observed strategy. This measure eliminates the variance present in the actual accuracy data due to the random location of the target and participant’s guessing the correct answer by chance. These data are summarised in Figure \@ref(fig:ProbExpAcc). We can see that in the symmetric condition, with the exception of one individual, our participants behaved in a way that gave rise to expected accuracies much closer to the counter-optimal strategy than optimal. Expected accuracy was higher in the biased condition, with four participants reaching optimal performance, with the rest ending up somewhere between the optimal and counter-optimal.
